### 2015-08-17 betaRegression/betaRegressionExample.Rmd
added files

---
title: "negativeBinomial"
author: "John Tipton"
date: "August 14, 2015"
output: html_document
---

# Write a beta GLM in STAN

```{r}
library(rstan)
model <- "
// beta Regression
data {
  int<lower=1> N;                   // sample size
  int<lower=1> p;                   // p predictors
  vector<lower=0,upper=1>[N] y;     // response 
  matrix[N, p] X;                   // predictor matrix
}
parameters {
  vector[p] theta;                  // reg coefficients
  real<lower=0> phi;                // dispersion parameter
}
transformed parameters{
  vector[p] beta;
  beta <- theta * 10;               // matt trick if desired
}
model {
  // model calculations
  vector[N] Xbeta;                  // linear predictor
  vector<lower=0,upper=1>[N] mu;    // transformed linear predictor
  vector<lower=0>[N] A;             // parameter for beta distn
  vector<lower=0>[N] B;             // parameter for beta distn
  Xbeta <- X * beta;
  for (i in 1:N) { 
    mu[i] <- inv_logit(Xbeta[i]);   
  }
  A <- mu * phi;
  B <- (1.0 - mu) * phi;
  // priors
  theta ~ normal(0, 1);   
  phi ~ cauchy(0, 5);               // different options for phi  
  //phi ~ inv_gamma(.001, .001);
  //phi ~ uniform(0, 500);          // put upper on phi if using this
  // likelihood
  y ~ beta(A, B);
}
"
write(model, file = "../stanModels/betaRegression-glm.stan")
sm <- stan_model("../stanModels/betaRegression-glm.stan")
```

## Simulate some Data

```{r}
library(betareg)
set.seed(123)
N = 80
beta <- c(2, -1, 1, -2, -1)
x0 <- rep(1, N)
x1 <- c(rep(1, N/2), rep(0, N/2)) ## Indicator for net size
x2 <- rep(c(rep(1, 10), rep(0, 30)), 2) ## Indicator for July
x3 <- rep(c(rep(0, 10), rep(1, 20), rep(0, 10)), 2) ## Indicator for August
# x4 <- rep(c(rep(0, 30), rep(1, 10)), 2) ## Indicator for September
x12 <- x1 * x2 ## Interaction between net size and July
x13 <- x1 * x3 ## Interaction between net size and August
# X <- cbind(x0, x1, x2, x3, x12, x13)
X <- cbind(x1, x2, x3, x12, x13)

mu = plogis(X%*%beta)  # add noise if desired + rnorm(N, sd=.01)
phi = 10
A = mu*phi
B = (1-mu)*phi
y = rbeta(N, A, B)
layout(matrix(c(1, 1, 2, 3), 2, 2, byrow=TRUE))
hist(y, "FD")
hist(y[x1 == 0], "FD")
hist(y[x1 == 1], "FD")
```

## Fit the Model

```{r}
m <- sampling(sm, data = list(N = N, y = y, X = X, p=dim(X)[2]),
  pars = c("beta", "phi"),
  iter = 5000, chains = 4)
```

## Model output
```{r}
m
```

## Plot traceplots
```{r}
rstan::traceplot(m, inc_warmup = FALSE)
```

## Plot fitted vs. simulated values

```{r}
e <- rstan::extract(m, pars = c("beta", "phi"))
true_pars <- c(beta = beta, phi = phi)
library(betareg)
brmod <- betareg(y ~ . -1, data = data.frame(y, X))
summary(brmod)
coefs <- c(brmod$coefficients$mean, brmod$coefficients$precision)
par(mfrow = c(2, 3))
for(i in 1:2) {
  if(i == 1) {
#     for(j in 1:6) {
    for(j in 1:5) {
      plot(density(e[[i]][,j]), main = names(true_pars)[i + j - 1])
      abline(v = true_pars[i + j - 1], lwd = 2, col = "grey", lty = 2)
      abline(v = coefs[i + j - 1], lwd = 3, col = "red")
    }
  } else if(i == 2) {
    plot(density(e[[i]]), main = names(true_pars)[i + j - 1])
    abline(v = true_pars[i + j - 1], lwd = 2, col = "grey", lty = 2)
    abline(v = coefs[i + j - 1], lwd = 3, col = "red")
  }
}
# legend("topright", legend = c("posterior", "true", "MASS::glm.nb"),
#   col = c("black", "grey", "red"), lty = c(1, 2, 1), lwd = c(1, 1.3, 3),
#   bty = "n")
```

```{r}
library(ggmcmc)
library(coda)
s <-  mcmc.list(lapply(1:ncol(m), function(x) mcmc(as.array(m)[,x,])))
S <- ggs(s)

idx <- order(colMeans(e$beta))

ggs_caterpillar(S, family = "beta") + 
  geom_point(aes(y=1:5, x=beta[idx]), col = "red")
```

```{r}
hist(y[X[, 1] == 0])
hist(y[X[, 1] == 1])
```

# model for later comparison

# simulate yRep as in BDA appendix C; could also do in generated quantities block of stan code
for (s in 1:nSim){
  betaS = betareg_sim$beta[s,]
  muS = plogis(X%*%betaS)
  A = muS*betareg_sim$phi[s]
  B = (1-muS)*betareg_sim$phi[s]
  yRep[s,] = rbeta(dat$N, A, B)
}
str(yRep)

# example, as in BDA p.144, code from appendix C; probably not very useful
par(mfrow=c(5,4), mar=c(4,4,2,2))
hist(y, xlab='', main='y', 'FD')
for (s in 1:19) hist(yRep[s,], xlab='', main=paste('yRep',s), 'FD')
layout(1)

# the following plots get at the same thing
library(ggplot2); library(reshape2)
gdat = data.frame(X,y)
gdatyRep = melt(yRep)

# individual observation distibutions; note the theme is my own (available on github), but otherwise you'll have to define yours.
ggplot(aes(x=y), xlim=c(.2,1), data=gdat) + 
  stat_density(geom='line', lwd=2, alpha=.2) + 
  geom_line(aes(x=value,  group=as.factor(Var2), col=as.factor(Var2)), stat='density', data=gdatyRep, show_guide=F, alpha=.15) +
  ggtheme

# whole response distributions, comparison with betareg fits
ggplot(aes(x=y), xlim=c(.2,1), data=gdat) + 
  geom_line(aes(x=value,  group=as.factor(Var1), col=as.factor(Var1)), stat='density', data=gdatyRep, show_guide=F, alpha=.05) +
  stat_density(geom='line', lwd=2, alpha=.2) + 
  stat_density(aes(x=brmod$fitted), geom='line', col='darkred', lwd=2, alpha=.2) +
  stat_density(aes(x=value), geom='line', col='#FF5500', lwd=2, alpha=.2, data=gdatyRep) +
  ggtheme

# Examine quantiles
sapply(list(yRep=gdatyRep$value, brFitted=brmod$fitted.values, y=y), quantile, p=c(0,.1,.25,.5,.625,.75,.8,.9,.95,1))