### 2015-08-24 sexRatioModel/sexRatioModel.Rmd

---
title: "Beta Regression"
author: "John Tipton"
date: "August 14, 2015"
output: html_document
---

## Note, there are propotions missing and proportions greater than 1

# Write a beta GLM in STAN

```{r}
library(rstan)
# model <- "
# // beta Regression
# data {
#   int<lower=1> N_obs;                   // sample size
#   int<lower=1> N_mis;                   // sample size
#   int<lower=1> p;                       // p predictors
#   vector<lower=0,upper=1>[N_obs] y_obs; // response 
#   matrix[N_obs, p] X_obs;               // predictor matrix
#   matrix[N_mis, p] X_mis;               // predictor matrix
# }
# parameters {
#   vector<lower=0,upper=1>[N_mis] y_mis; // missing response 
#   vector[p] theta;                      // reg coefficients
#   real<lower=0> phi;                    // dispersion parameter
# }
# transformed parameters{
#   vector[p] beta;
#   beta <- theta * 10;               // matt trick if desired
# }
# model {
#   // model calculations
#   vector[N_obs] Xbeta_obs;              // linear predictor
#   vector[N_mis] Xbeta_mis;              // linear predictor
#   vector<lower=0,upper=1>[N_obs] mu_obs;// transformed linear predictor
#   vector<lower=0,upper=1>[N_mis] mu_mis;// transformed linear predictor
#   vector<lower=0>[N_obs] A_obs;         // parameter for beta distn
#   vector<lower=0>[N_mis] A_mis;         // parameter for beta distn
#   vector<lower=0>[N_obs] B_obs;         // parameter for beta distn
#   vector<lower=0>[N_mis] B_mis;         // parameter for beta distn
#   Xbeta_obs <- X_obs * beta;
#   Xbeta_mis <- X_mis * beta;
#   for (i in 1:N_obs) { 
#     mu_obs[i] <- inv_logit(Xbeta_obs[i]);   
#   }
#   for (i in 1:N_mis) { 
#     mu_mis[i] <- inv_logit(Xbeta_mis[i]);   
#   }
#   A_obs <- mu_obs * phi;
#   A_mis <- mu_mis * phi;
#   B_obs<- (1.0 - mu_obs) * phi;
#   B_mis<- (1.0 - mu_mis) * phi;
#   // priors
#   theta ~ normal(0, 1);   
#   phi ~ cauchy(0, 5);               // different options for phi  
#   //phi ~ inv_gamma(.001, .001);
#   //phi ~ uniform(0, 500);          // put upper on phi if using this
#   // likelihood
#   y_obs ~ beta(A_obs, B_obs);
#   y_mis ~ beta(A_mis, B_mis);
# }
# "
# write(model, file = "~/mysis/stanModels/betaRegressionMissing-glm.stan")
sm <- stan_model("~/mysis/stanModels/betaRegression-glm.stan")
```

## Load the Data

```{r}
library(ggplot2)
set.seed(123)

fullData <- read.csv("~/mysis/data/lengthData.csv", skip=4)
substr(as.character(fullData$Gender), 0, 1)
fullData$Station

males <- rep(0, 80)
females <- rep(0, 80)
unknowns <- rep(0, 80)
size <- rep(0, 80)
date_tmp <- rep(0, 80)

idx <- 1
for(i in unique(fullData$Sample.Label)){
  tmp_data = subset(fullData, fullData$Sample.Label == i)
  tmp = substr(as.character(tmp_data$Gender), 0, 1)
  males[idx] <- sum(tmp == "M")
  females[idx] <- sum(tmp == "F")
  unknowns[idx] <- sum(tmp == "U")
  size[idx] <- tmp_data$Size..m.[1]
  date_tmp[idx] <- tmp_data$Date[1]
  idx <- idx + 1
}
prop <- females / (females + males)
prop_corrected <- (females + 0.5 * unknowns) / 
                  (females + males + unknowns)
prop[prop == 0] <- 0.01
prop_corrected[prop_corrected == 0] <- 0.01
## Correct for mislabeling
date_tmp[70] <- 4
date_tmp[72] <- 4
date <- rep(0, 80)
date[date_tmp == 1] <- 1
date[date_tmp == 2] <- 1
date[date_tmp == 3] <- 2
date[date_tmp == 4] <- 3

# data_one_way <- read.csv("~/mysis/data/ttestData.csv", skip = 3, nrows = 40, header = TRUE, stringsAsFactors=FALSE)

# ratio <- c(as.numeric(data_one_way$full.3),    ## proportion of females full net
#              as.numeric(data_one_way$half.3)) ## proportion of females half net
# N <- length(ratio)

# data_one_way$Sample.Date
## Set improper values to NA

date <- as.factor(date)
net <- as.factor(size)
data <- data.frame(prop=prop, date=date, net=net)
X <- model.matrix(~ net * date)
ggplot(data=data, aes(prop)) + geom_histogram() + facet_grid(net ~ .)

# date <- rep(as.factor(data_one_way$month), 2)
# net <- as.factor(c(rep("full", 40), rep("half", 40)))
# data <- data.frame(count=count, date=date, net=net)
# X <- model.matrix(~ net * date)
# ggplot(data=data, aes(prop)) + geom_histogram() + facet_grid(net ~ .)

```

## Fit the Model

```{r}
m <- sampling(sm, data = list(N=80, y=prop, X=X, p=dim(X)[2]),
  pars = c("beta", "phi"),
  iter = 5000, chains = 4)

m_corrected <- sampling(sm, data = list(N=80, y=prop_corrected, 
                                        X=X, p=dim(X)[2]),
  pars = c("beta", "phi"),
  iter = 5000, chains = 4)
```

## Model output
```{r}
m
```

## Plot traceplots
```{r}
rstan::traceplot(m, inc_warmup = FALSE)

rstan::traceplot(m_corrected, inc_warmup = FALSE)
```

## Plot fitted vs. simulated values

```{r}
e <- rstan::extract(m, pars = c("beta", "phi"))
library(betareg)
brmod <- betareg(y ~ . -1, data = data.frame(y=prop, X))
summary(brmod)
coefs <- c(brmod$coefficients$mean, brmod$coefficients$precision)
par(mfrow = c(3, 3))
for(i in 1:2) {
  if(i == 1) {
    for(j in 1:6) {
    # for(j in 1:5) {
      plot(density(e[[i]][,j]), main = paste("beta", j, sep=""))
      # abline(v = true_pars[i + j - 1], lwd = 2, col = "grey", lty = 2)
      abline(v = coefs[i + j - 1], lwd = 3, col = "red")
    }
  } else if(i == 2) {
    plot(density(e[[i]]), main = paste("phi"))
    # abline(v = true_pars[i + j - 1], lwd = 2, col = "grey", lty = 2)
    abline(v = coefs[i + j - 1], lwd = 3, col = "red")
  }
}

e_corrected <- rstan::extract(m_corrected, pars = c("beta", "phi"))
library(betareg)
brmod_corrected <- betareg(y ~ . -1, data = data.frame(y=prop_corrected, X))
summary(brmod_corrected)
coefs_corrected <- c(brmod_corrected$coefficients$mean, brmod_corrected$coefficients$precision)
par(mfrow = c(3, 3))
for(i in 1:2) {
  if(i == 1) {
    for(j in 1:6) {
    # for(j in 1:5) {
      plot(density(e_corrected[[i]][,j]), main = paste("beta", j, sep=""))
      # abline(v = true_pars[i + j - 1], lwd = 2, col = "grey", lty = 2)
      abline(v = coefs_corrected[i + j - 1], lwd = 3, col = "red")
    }
  } else if(i == 2) {
    plot(density(e_corrected[[i]]), main = paste("phi"))
    # abline(v = true_pars[i + j - 1], lwd = 2, col = "grey", lty = 2)
    abline(v = coefs_corrected[i + j - 1], lwd = 3, col = "red")
  }
}
```

```{r}
library(ggmcmc)
library(coda)
s <-  mcmc.list(lapply(1:ncol(m), function(x) mcmc(as.array(m)[,x,])))
S <- ggs(s)

ggs_caterpillar(S, family = "beta")

s_corrected <-  mcmc.list(lapply(1:ncol(m_corrected), function(x) mcmc(as.array(m_corrected)[,x,])))
S_corrected <- ggs(s_corrected)

ggs_caterpillar(S_corrected, family = "beta") 
```

# model for later comparison

# simulate yRep as in BDA appendix C; could also do in generated quantities block of stan code
for (s in 1:nSim){
  betaS = betareg_sim$beta[s,]
  muS = plogis(X%*%betaS)
  A = muS*betareg_sim$phi[s]
  B = (1-muS)*betareg_sim$phi[s]
  yRep[s,] = rbeta(dat$N, A, B)
}
str(yRep)

# example, as in BDA p.144, code from appendix C; probably not very useful
par(mfrow=c(5,4), mar=c(4,4,2,2))
hist(y, xlab='', main='y', 'FD')
for (s in 1:19) hist(yRep[s,], xlab='', main=paste('yRep',s), 'FD')
layout(1)

# the following plots get at the same thing
library(ggplot2); library(reshape2)
gdat = data.frame(X,y)
gdatyRep = melt(yRep)

# individual observation distibutions; note the theme is my own (available on github), but otherwise you'll have to define yours.
ggplot(aes(x=y), xlim=c(.2,1), data=gdat) + 
  stat_density(geom='line', lwd=2, alpha=.2) + 
  geom_line(aes(x=value,  group=as.factor(Var2), col=as.factor(Var2)), stat='density', data=gdatyRep, show_guide=F, alpha=.15) +
  ggtheme

# whole response distributions, comparison with betareg fits
ggplot(aes(x=y), xlim=c(.2,1), data=gdat) + 
  geom_line(aes(x=value,  group=as.factor(Var1), col=as.factor(Var1)), stat='density', data=gdatyRep, show_guide=F, alpha=.05) +
  stat_density(geom='line', lwd=2, alpha=.2) + 
  stat_density(aes(x=brmod$fitted), geom='line', col='darkred', lwd=2, alpha=.2) +
  stat_density(aes(x=value), geom='line', col='#FF5500', lwd=2, alpha=.2, data=gdatyRep) +
  ggtheme

# Examine quantiles
sapply(list(yRep=gdatyRep$value, brFitted=brmod$fitted.values, y=y), quantile, p=c(0,.1,.25,.5,.625,.75,.8,.9,.95,1))