### 2015-10-12 juvenileRatioModel/juvenileRatioModel.Rmd

---
title: "Beta Regression"
author: "John Tipton"
date: "August 14, 2015"
output: html_document
---
# Write a beta GLM in STAN

```{r}
library(rstan)
sm <- stan_model("~/mysis/stanModels/betaRegression-glm.stan")
```

## Load the Data

```{r}
library(ggplot2)
set.seed(123)
data_one_way <- read.csv("~/mysis/data/ttestData.csv", skip = 3, nrows = 40, header = TRUE, stringsAsFactors=FALSE)

prop <- c(as.numeric(data_one_way$full.2),    ## proportion of females full net
             as.numeric(data_one_way$half.2)) ## proportion of females half net
N <- length(prop)
## Set improper values to NA

date <- rep(as.factor(data_one_way$month), 2)
net <- as.factor(c(rep("full", 40), rep("half", 40)))
data <- data.frame(prop=prop, date=date, net=net)
X <- model.matrix(~ net * date)
ggplot(data=data, aes(prop)) + geom_histogram() + facet_grid(net ~ .)
```

## Fit the Model

```{r}
m <- sampling(sm, data = list(N=N, y = prop, X = X, p=dim(X)[2]),
  pars = c("beta", "phi"),
  iter = 5000, chains = 4)
```

## Model output
```{r}
m
```

## Plot traceplots
```{r}
rstan::traceplot(m, inc_warmup = FALSE)
```

## Plot fitted vs. simulated values

```{r}
e <- rstan::extract(m, pars = c("beta", "phi"))
library(betareg)
brmod <- betareg(prop ~ . -1, data = data.frame(prop, X))
summary(brmod)
coefs <- c(brmod$coefficients$mean, brmod$coefficients$precision)
par(mfrow = c(3, 3))
for(i in 1:2) {
  if(i == 1) {
    for(j in 1:6) {
    # for(j in 1:5) {
      plot(density(e[[i]][,j]), main = paste("beta", j, sep=""))
      # abline(v = true_pars[i + j - 1], lwd = 2, col = "grey", lty = 2)
      abline(v = coefs[i + j - 1], lwd = 3, col = "red")
    }
  } else if(i == 2) {
    plot(density(e[[i]]), main = "phi")
    # abline(v = true_pars[i + j - 1], lwd = 2, col = "grey", lty = 2)
    abline(v = coefs[i + j - 1], lwd = 3, col = "red")
  }
}
# legend("topright", legend = c("posterior", "true", "MASS::glm.nb"),
#   col = c("black", "grey", "red"), lty = c(1, 2, 1), lwd = c(1, 1.3, 3),
#   bty = "n")
```

```{r}
library(ggmcmc)
library(coda)
s <-  mcmc.list(lapply(1:ncol(m), function(x) mcmc(as.array(m)[,x,])))
S <- ggs(s)

ggs_caterpillar(S, family = "beta")

```

# model for later comparison

# simulate yRep as in BDA appendix C; could also do in generated quantities block of stan code
for (s in 1:nSim){
  betaS = betareg_sim$beta[s,]
  muS = plogis(X%*%betaS)
  A = muS*betareg_sim$phi[s]
  B = (1-muS)*betareg_sim$phi[s]
  yRep[s,] = rbeta(dat$N, A, B)
}
str(yRep)

# example, as in BDA p.144, code from appendix C; probably not very useful
par(mfrow=c(5,4), mar=c(4,4,2,2))
hist(y, xlab='', main='y', 'FD')
for (s in 1:19) hist(yRep[s,], xlab='', main=paste('yRep',s), 'FD')
layout(1)

# the following plots get at the same thing
library(ggplot2); library(reshape2)
gdat = data.frame(X,y)
gdatyRep = melt(yRep)

# individual observation distibutions; note the theme is my own (available on github), but otherwise you'll have to define yours.
ggplot(aes(x=y), xlim=c(.2,1), data=gdat) + 
  stat_density(geom='line', lwd=2, alpha=.2) + 
  geom_line(aes(x=value,  group=as.factor(Var2), col=as.factor(Var2)), stat='density', data=gdatyRep, show_guide=F, alpha=.15) +
  ggtheme

# whole response distributions, comparison with betareg fits
ggplot(aes(x=y), xlim=c(.2,1), data=gdat) + 
  geom_line(aes(x=value,  group=as.factor(Var1), col=as.factor(Var1)), stat='density', data=gdatyRep, show_guide=F, alpha=.05) +
  stat_density(geom='line', lwd=2, alpha=.2) + 
  stat_density(aes(x=brmod$fitted), geom='line', col='darkred', lwd=2, alpha=.2) +
  stat_density(aes(x=value), geom='line', col='#FF5500', lwd=2, alpha=.2, data=gdatyRep) +
  ggtheme

# Examine quantiles
sapply(list(yRep=gdatyRep$value, brFitted=brmod$fitted.values, y=y), quantile, p=c(0,.1,.25,.5,.625,.75,.8,.9,.95,1))